---
title: "Relative Abundances"
output: html_document
date: "2024-09-20"
---
```{r setup, include = FALSE, fig.path = "custom/path", warning = FALSE, message = FALSE}

options (width = 100)
knitr::opts_chunk$set(warning = FALSE, echo = TRUE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234

```
```



```{r load libraries, echo=FALSE, message=FALSE, warning=FALSE}

suppressPackageStartupMessages({
  library(ExploreSCdataSeurat3)
  library(runSeurat3)
  library(Seurat)
  library(ggpubr)
  library(pheatmap)
  library(SingleCellExperiment)
  library(dplyr)
  library(tidyverse)
  library(viridis)
  library(muscat)
  library(circlize)
  library(destiny)
  library(scater)
  library(metap)
  library(multtest)
  library(clusterProfiler)
  library(org.Hs.eg.db)
  library(msigdbr)
  library(enrichplot)
  library(DOSE)
  library(grid)
  library(gridExtra)
  library(ggupset)
  library(VennDiagram)
  library(NCmisc)
  library(RColorBrewer)
  library(EnhancedVolcano)
  library(textshaping)
  library(wordcloud)
  library(pathview)
  library(workflowr)
})

```


```{r}

##load merged file 
seuratM <- readRDS("~/Desktop/Emily/Myocarditis snRNAseq/data/01 Merged Seurat files/Myocarditis_RDS_04.10.2024.rds")

table(seuratM$dataset)
table(seuratM$RNA_snn_res.0.25)
table(seuratM$orig.ident)

```







```{r}


seuratM$clusterName <- "clusterName"
Idents(seuratM) <- seuratM$clusterName
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "0" )] <- "Fb1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "1" )] <- "BEC1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "2" )] <- "PeriFb1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "3" )] <- "Cardiomyocyte"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "4" )] <- "MCP1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "5" )] <- "TC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "6" )] <- "MCP2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "7" )] <- "BEC2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "8" )] <- "VSMC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "9" )] <- "NC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "10" )] <- "INT1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "11" )] <- "Fb2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "12" )] <- "Fb3"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "13" )] <- "PeriFb2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "14" )] <- "AdipoC"

seuratM$clusterName <- factor(seuratM$clusterName, levels =c("TC", "MCP1","MCP2","Cardiomyocyte","Fb1","Fb2","Fb3", "PeriFb1", "PeriFb2","VSMC", "INT1","NC", "BEC1","BEC2", "AdipoC"))

##set color vectors

colclusterName <- c("#dd1c77", "#fde0dd", "#B45B5C", "#D53E4F", "#f4a582", "#FEE08B", "#feb24c", "#01665e","#66C2A5", "#c7eae5", "#67001f", "#BEAED4", "#355C7D","#3288BD","#8c510a")
                    
names(colclusterName) <- c("TC", "MCP1","MCP2","Cardiomyocyte","Fb1","Fb2","Fb3", "PeriFb1", "PeriFb2","VSMC", "INT1","NC", "BEC1","BEC2", "AdipoC")


colpat_sub<- c("#dfc27d","#BE3144","#355C7D", "#779d8d")

coldiseaseCond <- colpat_sub
names(coldiseaseCond) <- c("Healthy", "AM", "CMP", "Sarcoidosis")
order_coldiseaseCond <- c("Healthy", "AM", "CMP", "Sarcoidosis")

head(seuratM@meta.data)

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap", pt.size = 0.1, cols = colclusterName)

### Order of Conditions
Idents(seuratM) <- seuratM$pat_sub
DimPlot(seuratM, reduction = "umap", pt.size = 0.1, group.by = "pat_sub", cols = coldiseaseCond)

```

```{r Relative abundance per disease condition}

datList <- NULL
for(con in unique(seuratM$pat_sub)){
  seuratSub <- subset(seuratM, pat_sub==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(diseaseCond=con)
  datList[[con]] <- dat_con
  
}

dat_all <- do.call("rbind", datList)

order_patsub <- c("Healthy", "AM", "CMP", "Sarcoidosis")

ggbarplot(dat_all, x = "diseaseCond", y = "percent", fill = "Var1", palette = colclusterName, legend = "right", legend.titel = "cluster", xlab = "condition", ylab = "frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = order_patsub)




```

```{r Relative abundance per patient}

datList <- NULL
for(con in unique(seuratM$patient)){
  seuratSub <- subset(seuratM, patient==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(patient=con)
  datList[[con]] <- dat_con
}


dat_all <- do.call("rbind", datList)

# Order 
patient_order <- c("Healthy 01", "Healthy 02", "Healthy 03", "Healthy 04", "Healthy 05", 
                   "Healthy 06", "Healthy 07", "Healthy 08", "Healthy 09", "Healthy 10", 
                   "AM 01", "AM 02", "AM 03", "AM 04", "AM 05", "AM 06", "AM 07", 
                   "AM 08", "AM 09", "AM 10", "AM 11", "AM 12", "AM 13", "AM 14", 
                   "AM 15", "AM 16", "AM 17", "AM 18", "CMP 01", "CMP 02", "CMP 03", 
                   "CMP 04", "CMP 05", "CMP 06", "CMP 07", "CMP 08", "CMP 09", 
                   "CMP 10", "CMP 11", "CMP 12", "CMP 13", "Sarcoidosis 01", "Sarcoidosis 02")

#View(dat_all)

#write.table(dat_all, file= "/Users/immbio/Desktop/Emily/Myocarditis snRNAseq/docs/Markers_dat_all_clusters_frequency_relative abundance_V2", sep = "\t",quote=F,row.names=T,col.names=T)


dat_all$patient <- factor(dat_all$patient, levels = patient_order)

## plot abundance
ggbarplot(dat_all, x= "patient", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency", palette = colclusterName )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## Change order

# Subset TC cluster
dat_tc <- dat_all[dat_all$Var1 == "TC", ]

# T cells in descending order
dat_all$patient <- factor(dat_all$patient, levels = dat_tc$patient[order(-dat_tc$percent)])

ggbarplot(dat_all, x= "patient", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency", palette = colclusterName )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## Inflammatory Macrophages MCP2 in descending order
dat_mcp2 <- dat_all[dat_all$Var1 == "MCP2", ]

dat_all$patient <- factor(dat_all$patient, levels = dat_mcp2$patient[order(-dat_mcp2$percent)])

ggbarplot(dat_all, x= "patient", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency", palette = colclusterName )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



```









```{r Heatmap with marker genes, fig.height = 20}

## Create Heatmap

Idents(seuratM) <- seuratM$clusterName
seuratM$clusterName <- factor(seuratM$clusterName, levels =c("Cardiomyocyte","Fb1","Fb2","Fb3", "INT1", "PeriFb1", "PeriFb2", "VSMC","NC", "BEC1","BEC2", "AdipoC", "MCP1","MCP2","TC"))

genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

selGenes <- data.frame(geneID=rev(c("TTN", "MYBPC3", "RYR2", "NEBL", "TNNT2", "CMYA5", "COL6A3", "DCN",  "FBN1", "C7", "PDGFRA", "CDH19", "PDGFRB", "BMP4", "GREM1", "ITGA7","RGS5", "NOTCH3", "MYH11", "ACTA2","PECAM1", "VWF", "EGFL7", "POSTN", "ITGA10", "CDH11","CCL21", "FLT4", "NRXN1", "ANK3", "PTPRZ1", "PLIN1", "GPAM", "CD163", "LYVE1", "MRC1", "SIGLEC1", "STAB1", "CSF1R","MERTK", "HLA-DQB1","CCR2", "CD79B","JCHAIN", "IL7R", "CD3E", "PTPRC", "CD2"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5") 


avgHeatmap <- function(seuratM, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seuratM)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seuratM)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seuratM)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      celltype=colVec)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
  ordVec <- levels(seuratM)
  ordVec <- c(c("Cardiomyocyte","Fb1","Fb2","Fb3", "INT1", "PeriFb1", "PeriFb2", "VSMC","NC", "BEC1","BEC2", "AdipoC", "MCP1","MCP2","TC"))
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0,cluster_rows = cr, 
         cluster_cols = cc, color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

genesPlot <- data.frame(gene=c("TTN", "MYBPC3", "RYR2", "NEBL", "TNNT2", "CMYA5", "COL6A3", "DCN",  "FBN1", "C7", "PDGFRA", "CDH19", "PDGFRB", "BMP4", "GREM1", "ITGA7","RGS5", "NOTCH3", "MYH11", "ACTA2","PECAM1", "VWF", "EGFL7", "POSTN", "ITGA10", "CDH11","CCL21", "FLT4", "NRXN1", "ANK3", "PTPRZ1", "PLIN1", "GPAM", "CD163", "LYVE1", "MRC1", "SIGLEC1", "STAB1", "CSF1R","MERTK", "HLA-DQB1", "CCR2", "CD79B","JCHAIN", "IL7R","CD3E", "PTPRC", "CD2"))

levels(seuratM)


colVec <- colclusterName

# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")

avgHeatmap(seuratM, selGenes = genesPlot, colVecIdent = colVec)


```


```{r}

date()
sessionInfo()

```










`
